---
output:
  word_document: default
  html_document: default
---
#Project 1-Data Exploration, Preparation, and Visualization

##BAN502 Predictive Analytics

###Susan Wiggins

Libraries needed for Project 1 Data Exploration:

```{r}

options(tidyverse.quiet = TRUE)
library(tidyverse)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(leaps) #best subset selection
library(MASS) #access to forward and backward selection algorithms
library(weathermetrics)
library(caret) #for splitting functions

```

Reading in the data set:

```{r}

library(readr)
rain <- read_csv("rain.csv")
View(rain)


```

###Structure and summary and Visualizations for understanding of variables in the data set:




```{r}


str(rain)
summary(rain)

```

```{r}



```



Converting variables to factors:

```{r}

rain = rain %>% mutate(RainTomorrow = as_factor(as.character(RainTomorrow)))%>% 
mutate(RainTomorrow = fct_recode(RainTomorrow, "Yes" = "1", "No" = "0"))

glimpse(rain)


rain = rain %>% mutate(Cloud9am = as_factor(as.character(Cloud9am)))%>% 
mutate(Cloud9am = fct_recode(Cloud9am, "clear sky" = "0", "few clouds" = "1",  "2", "scattered clouds" = "3", "4","broken" = "5", "6","7", "completely overcast" = "8"))


glimpse(rain)


rain = rain %>% mutate(Cloud3pm = as_factor(as.character(Cloud3pm)))%>% 
mutate(Cloud3pm = fct_recode(Cloud3pm, "clear sky" = "0", "few clouds" = "1",  "2", "scattered clouds" = "3", "4","broken" = "5", "6","7", "completely overcast" = "8"))


glimpse(rain)


rain = rain %>% mutate(RainToday = as_factor(as.character(RainToday)))%>% 
mutate(RainToday = fct_recode(RainToday, "Yes" = "1", "No" = "0"))

glimpse(rain)



```

Converting Celsius to Farhenheit since United States looks at this scale for 
Temperature readings:

```{r}



data(rain)
rain$MinTempF <- convert_temperature(rain$MinTemp,
   old_metric = "celsius", new_metric = "fahrenheit")

rain$MaxTempF <- convert_temperature(rain$MaxTemp,
   old_metric = "celsius", new_metric = "fahrenheit")

rain$Temp9amF <- convert_temperature(rain$Temp9am,
   old_metric = "celsius", new_metric = "fahrenheit")

rain$Temp3pmF <- convert_temperature(rain$Temp3pm,
   old_metric = "celsius", new_metric = "fahrenheit")


rain$Rainfall_inches <- convert_precip(rain$Rainfall,
  old_metric = "mm", new_metric = "inches", round = NULL)

str(rain)
summary(rain)




```

Difference in the MaxTemperture and the Minimum Temperature in the data set along with a conversion on the temperature scale:

```{r}

rain <- mutate(rain, Temperture_Change = MaxTemp - MinTemp)


rain$Temperture_ChangeF <- convert_temperature(rain$Temperture_Change,
   old_metric = "celsius", new_metric = "fahrenheit")


rain$Cloud3pm[rain$Cloud3pm==""] = NA #convert blanks in cabin to NA
rain$Cloud9am[rain$Cloud9am==""] = NA #convert blanks in cabin to NA

str(rain)
summary(rain)



```



```{r}

```



Visualization of Variables in Rain dataset:

```{r}

ggplot(rain, aes(x= Date, fill = RainTomorrow)) + geom_bar() + theme_bw()



```


```{r}

ggplot(rain, aes(x= MinTemp, fill = RainTomorrow)) + geom_bar() + theme_bw()



```


```{r}


ggplot(rain, aes(x= MaxTemp, fill = RainTomorrow)) + geom_bar() + theme_bw()

```

```{r}

ggplot(rain, aes(x= Rainfall, fill = RainTomorrow)) + geom_bar() + 
  theme_bw() + scale_x_continuous(limit = c(0, 25)) +
   scale_y_continuous(limit = c(0, 1750))
```

```{r}


ggplot(rain, aes(x= Cloud9am, fill = RainTomorrow)) + geom_bar() + 
  theme_bw() + theme(axis.text.x = element_text(angle=60, hjust=1))



```


```{r}


ggplot(rain, aes(x= Cloud3pm, fill = RainTomorrow)) + geom_bar() + 
  theme_bw() + theme(axis.text.x = element_text(angle=60, hjust=1)) 


```


```{r}



ggplot(rain, aes(x= RainToday , fill = RainTomorrow)) + geom_bar() + 
  theme_bw() 



```

```{r}


t_rain = table(rain$RainTomorrow, rain$RainToday) #create a table object
prop.table(t_rain, margin = 2 ) #crosstab with proportions



```


```{r}


ggplot(rain, aes(x= Temp9am , fill = RainTomorrow)) + geom_bar() + 
  theme_bw() 


```



```{r}


ggplot(rain, aes(x= Temp3pm , fill = RainTomorrow)) + geom_bar() + 
  theme_bw() 




```


```{r}



ggplot(rain, aes(x= WindGustDir , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 




```



```{r}


ggplot(rain, aes(x= WindGustSpeed , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 


```



```{r}

ggplot(rain, aes(x= WindDir9am , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 




```

```{r}


ggplot(rain, aes(x= WindDir3pm, fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 


```

```{r}

ggplot(rain, aes(x= WindSpeed9am , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 




```


```{r}


ggplot(rain, aes(x= WindSpeed3pm , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 


```


```{r}

ggplot(rain, aes(x= Humidity9am , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 




```


```{r}

ggplot(rain, aes(x= Humidity3pm , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 



```


```{r}

ggplot(rain, aes(x= Pressure9am , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 



```


```{r}


ggplot(rain, aes(x= Pressure3pm , fill = RainTomorrow)) + geom_bar() + 
  theme_bw()  + theme(axis.text.x = element_text(angle=60, hjust=1)) 


```


```{r}




rain$Temperture_Change %>%  boxplot(main="Box Plot of Temperature Change", ylab="TempertureChange", col = "grey")



```


```{r}


ggplot(rain, aes(x= MinTempF, fill = RainTomorrow)) + geom_bar() + theme_bw()



```


```{r}



ggplot(rain, aes(x= MaxTempF, fill = RainTomorrow)) + geom_bar() + theme_bw()


```



```{r}


ggplot(rain, aes(x= Temp9amF, fill = RainTomorrow)) + geom_bar() + theme_bw()


```



```{r}


ggplot(rain, aes(x= Temp3pmF, fill = RainTomorrow)) + geom_bar() + theme_bw()



```

###Identify Missingness of data and appropriately deal with missingness:

```{r}

vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)


```


Removal of Cloud9am and Cloud3pm due to such a large degree of data missing around over 60% missing data points. 

Column-wise deletion of the "Cloud9am" and "Cloud3pm" variable, not sure why this was not removed. Next, Letting use the "mice" package to do imputation instead.


```{r}

#rain = rain %>% select(-Cloud9am) 
#vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)

#rain = rain %>% select(-Cloud3pm) 
#vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)

rain_subset = rain %>% dplyr:: select(Date, MinTempF, MaxTempF, Rainfall_inches, RainTomorrow, RainToday, Temperture_ChangeF, Humidity9am, Humidity3pm)
head(rain)

imp_age = mice(rain_subset, m=1, method='pmm', printFlag=FALSE)
summary(imp_age)


```


```{r}

densityplot(imp_age, ~Temperture_ChangeF) #red imputed, blue original


```

```{r}

densityplot(imp_age, ~MinTempF) #red imputed, blue original


```


```{r}

densityplot(imp_age, ~MaxTempF) #red imputed, blue original


```



```{r}

densityplot(imp_age, ~RainToday) #red imputed, blue original



```


```{r}

densityplot(imp_age, ~Rainfall_inches) #red imputed, blue original


```


```{r}

densityplot(imp_age, ~Humidity9am) #red imputed, blue original

```


```{r}

densityplot(imp_age, ~Humidity3pm) #red imputed, blue original

```



```{r}


```



Merge the imputed values into our data frame 

```{r}

rain_complete = complete(imp_age) 
summary(rain_complete)

```



```{r}



set.seed(12345)
train.rows = createDataPartition(y = rain$RainTomorrow, p=0.7, list = FALSE) #70% in training
train = rain[train.rows,] 
test = rain[-train.rows,]



```


